### 2018-01-19 related_work.tex
Diez: Added related word and an article

@article{doi:10.1080/01449290903121386,
author = { L.   Tam  and  M.   Glassman  and  M.   Vandenwauver },
title = {The psychology of password management: a tradeoff between security and convenience},
journal = {Behaviour \& Information Technology},
volume = {29},
number = {3},
pages = {233-244},
year  = {2010},
publisher = {Taylor & Francis},
doi = {10.1080/01449290903121386},

URL = { 
        https://doi.org/10.1080/01449290903121386

},
eprint = { 
        https://doi.org/10.1080/01449290903121386

}

}

@article{HUANG20111292,
title = "Using one-time passwords to prevent password phishing attacks",
journal = "Journal of Network and Computer Applications",
volume = "34",
number = "4",
pages = "1292 - 1301",
year = "2011",
note = "Advanced Topics in Cloud Computing",
issn = "1084-8045",
doi = "https://doi.org/10.1016/j.jnca.2011.02.004",
url = "http://www.sciencedirect.com/science/article/pii/S1084804511000427",
author = "Chun-Ying Huang and Shang-Pin Ma and Kuan-Ta Chen",
keywords = "Anti-phishing, Identity management, In-band password delivery, One-time password, Web security"
}

http://www.computerweekly.com/answer/Techniques-for-preventing-a-brute-force-login-attack

https://www.researchgate.net/publication/299645572_Data_protection_and_Brute_Force_attack